Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2020-04-27 manuscript/manuscript.tex
ms updates

During this burnin, sites were either marked as neutral (with mutation rate
$\mu_\mathrm{neutral} = 10^8$ per gamete per generation) or contributed to the
trait's value (with mutation rate $\mu_\mathrm{trait}$), but were not selected
on until generation $10N + 5$ (the five generations after burnin serve as a
neutral control). The trait mutation rate, $\mu_\mathrm{trait}$ was set by
targeting a particular architecture, the number of selected sites, $L$. Each
site contributing to the trait's value was randomly chosen to have effect size
$\pm \alpha$ with equal probability, where $\alpha$ was set to target a
particular additive genetic variance for the trait, $V_A$, for the target
number of selected sites $L$.

@@

  \caption{The $G(t)$ trajectories of 50 replicate simulations, across
    different trait architectures ($L$ is the target number of sites affecting
    the trait's value, and $V_A$ is the target trait additive genetic
    variance).  The red line is the mean trajectory across all replicate
  simulations. Like Supplementary Figure \ref{suppfig:sim-expfit-covs}, the
onset of selection is five generations after the $10N$ generation burnin; this
is evident by the initial flat section of the $G(t)$ trajectory.}

@@

@@

Overall, our simulations confirm our hypotheses; see Supplementary Material
Figure \ref{suppfig:convergence-corrs}. We also find that in simulations where
we target a monogenic genetic architecture (i.e. the target number of
trait-affecting loci is $L=1$), the convergence correlations are generally much
weaker than those under a polygenic architecture. However, this effect is
mediated by the line population size; the difference in convergence correlation
between $L = 1$ and $L = 1000$ are more dissimilar when the line population
sizes are larger (compare the first column, last two rows). Overall, we find
that the convergence correlation is affected by both genetic architecture and
the size of the selected population lines.

We also wanted to test whether our we see similar convergence correlations
under a more complicated, but more realistic fitness model: Gaussian
stabilizing selection. To this end we simulated both sudden optima shifts of
$\mu_\text{sudden} = \{0.1\sigma, 0.5\sigma, 1\sigma\}$ and very graduate
optima shifts of $\mu_\text{gradual} = \{\0.001\sigma, 0.01\sigma\}$ per
generation using the same two population simulation scheme described above. We
show these convergence correlations for the sudden optima shifts in
Supplementary Figure \ref{suppfig:convergence-corrs-gss}. We found that for
slow moving optima shifts, the convergence correlations are generally too weak
to be distinguished from noise reliably, so these results are excluded (but see
Supplementary Materials Section XXX).

>>>>

During this burnin, sites were either marked as neutral (with mutation rate
$\mu_\mathrm{neutral} = 10^{-8}$ per gamete per generation) or contributed to
the trait's value (with mutation rate $\mu_\mathrm{trait}$), but were not
selected until generation $10N + 5$ (the five generations after burnin serve
as a neutral control). The trait mutation rate, $\mu_\mathrm{trait}$ was set by
targeting a particular architecture, the number of selected sites, $L$. Each
site contributing to the trait's value was randomly chosen to have effect size
$\pm \alpha$ with equal probability, where $\alpha$ was set to target a
particular additive genetic variance for the trait, $V_A$, for the target
number of selected sites $L$.

@@

  \caption{The $G(t)$ trajectories of 50 replicate simulations, across
    different trait architectures ($L$ is the target number of sites affecting
    the trait's value, and $V_A$ is the target trait additive genetic
    variance).  The red line is the mean trajectory across all replicate
  simulations. Like Supplementary Figure \ref{suppfig:sim-expfit-covs}, the
onset of selection is five generations after the $10N$ generation burnin; this
is evident by the initial flat period of the $G(t)$ trajectory.}

@@

Additionally, we wanted to ensure that our temporal covariances and $G(t)$
trajectories were robust to more complicated, but realistic fitness models. To
this end, we simulated Gaussian stabilizing selection on a trait during burnin,
with two optima shift routines: (1) sudden optima shifts of $\mu_\text{sudden}
= \{0.1\sigma, 0.5\sigma, 1\sigma\}$, and (2) very graduate optima shifts of
$\mu_\text{gradual} = \{0.001\sigma, 0.01\sigma\}$ per generation using the
same two population simulation scheme described above. We used a polygenic
architecture for these simulations ($L = 1000$ trait affecting loci), and the
optima shift began at five generations after a $10N$ generation.

\begin{figure}[!ht]
  \centering
  \includegraphics[width=\textwidth]{figures/gss-covs.pdf}

  \label{gss-covs}
\end{figure}

\begin{figure}[!ht]
  \centering
  \includegraphics[width=\textwidth]{figures/gss-G.pdf}

  \label{gss-G}
\end{figure}

@@

Overall, our simulations confirm our hypotheses; see Supplementary Material
Figure \ref{suppfig:convergence-corrs}. We also find that in simulations where
we target a monogenic genetic architecture (i.e. the target number of
trait-affecting loci is $L=1$), the convergence correlations are generally much
weaker than those under a polygenic architecture. However, this effect is
mediated by the line population size; the difference in convergence correlation
between $L = 1$ and $L = 1000$ are more dissimilar when the line population
sizes are larger (compare the first column, last two rows). Like the
convergence correlations calculated on the \textcite{Barghi2019-qy} data, we
find in simulations convergence correlations decay through time. Overall, we
find that the convergence correlation is affected by both genetic architecture
and the size of the selected population lines.

We also wanted to test whether our we see similar convergence correlations
under Gaussian stabilizing selection. XXX We show these convergence
correlations for the sudden optima shifts in Supplementary Figure
\ref{suppfig:convergence-corrs-gss}. We found that for slow moving optima
shifts, the convergence correlations are generally too weak to be distinguished
from noise reliably, so these results are excluded (but see Supplementary
Materials Section XXX).
